
---
title: "PCA Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document demonstrates how to perform Principal Component Analysis (PCA) on a dataset. PCA is a technique used to reduce dimensionality while preserving as much variance as possible.

---

## Step 1: Load the Data

```{r load-data}
# Load necessary libraries
library(tidyverse)

# Load the dataset (assuming it's in the same folder as this Rmd file)
data <- read.csv("pca_data.csv")

# Quick look at the data
head(data)
summary(data)
```

---

## Step 2: Perform PCA

```{r perform-pca}
# Standardize the data
data_scaled <- scale(data)

# Perform PCA
pca_model <- prcomp(data_scaled, center = TRUE, scale. = TRUE)

# Summary of PCA
summary(pca_model)

# PCA loadings
pca_model$rotation
```

---

## Step 3: Visualize the Results

```{r visualize-pca}
# Load ggplot2 for visualization
library(ggplot2)

# Scree plot
pca_var <- pca_model$sdev^2
pca_var_explained <- pca_var / sum(pca_var)

ggplot(data.frame(PC = 1:length(pca_var), Variance = pca_var_explained), aes(x = PC, y = Variance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Scree Plot", x = "Principal Component", y = "Proportion of Variance Explained") +
  theme_minimal()

# Biplot
biplot(pca_model, scale = 0)
```

---

## Conclusion

This document demonstrates PCA for dimensionality reduction and visualization. You can extend this workflow to analyze more complex datasets or compare PCA results across different datasets.
